Deep Reinforcement Learning for Conservation Decisions

### 2021-05-05 manuscript/manuscript.Rmd
:pen:

<!-- Environmental problems, ML -->
Advances in both available data and computing power have begun to open the door to a greater role for machine learning (ML) in addressing some of our planet’s most pressing environmental problems, such as the growing frequency and intensity of wildfire [@wildfire], over-exploited fisheries [@worm2006], declining biodiversity [@], and zoonotic pandemics [@covid].
But will ML approaches really help us tackle a changing planet?
Applications of ML in ecology have begun to illustrate the promise of two methods of ML: *supervised learning* [@joseph2020] and *unsupervised learning* [@unsupervised], but have so far largely overlooked the third and possibly most promising approach in the ML triad: *reinforcement learning* (RL).
Three features distinguish RL from other ML methods in ways that are particularly well suited to addressing global ecological change issues:

@@

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the problems considered by RL research have largely been drawn from examples in robotic movement and games like Go and Starcraft [@lillicrap; @alphazero; @alphastar]
Complex environmental problems share many similarities to these tasks and games: the need to plan many moves ahead given a large number of possible outcomes, to account for uncertainty and to respond with contingency to the unexpected.
RL agents develop their strategies by interacting with simulators, a practice that should not be unsettling to ecologists since learning from simulators is a common practice.
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or global circulation models (GCMs) in climate [@gcm] are already used to explore scenarios and inform ecosystem management. 
Even simpler simulators are frequently employed throughout ecology to guide understanding and response.
<!-- Kari: ick, jargon heavy.  Need to say: some existing methods, but limited to simple problems or simple solutions -->
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of ecological models; the so called "curse of dimensionality" makes problems with a large number of states or actions intractable by conventional methods such as stochastic dynamic programming (SDP) and approximate dynamic programming (ADP) [@Polasky2011; @Marescot2013].
Neural-network-based RL techniques, referred to as *deep RL*, have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.
We believe that the application of RL methods to simulations of ecological processes will open the door to a new paradigm for tackling tough environmental change problems.
We also believe that such a transformation could bring risks and pitfalls comparable and varied as its solutions.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{fig1.pdf}
\caption{Reinforcement Learning focuses on the repeated feedback of agent interacting with an \emph{environment}. The agent performs an action which influences the environment's future \emph{state}.  The agent receives an observation of the new state and reward from the environment.}
\end{figure}

@@

I think we can lean more heavily both in references to and paralleling the presentation of some of the existing background material.

@@

Meanwhile, the state-action-value or quality function, finds the expected return given a state and action,

$$
  Q^\pi(s_t, a_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t, a_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

@@

Should we mention:

- Temporal Differencing, RIENFORCE
- SARSA, 
- Double DQN?  
Can we add HER (since delayed reward case may be particularly interesting)
Should we do anything to better reflect the 'family tree' here? (e.g. DQN -> DDPG -> TD3)

>>>>

<!-- Environmental problems, ML -->
Advances in both available data and computing power have begun to open the door to a greater role for machine learning (ML) in addressing some of our planet’s most pressing environmental problems, such as the growing frequency and intensity of wildfire [@wildfire], over-exploited fisheries [@worm2006], declining biodiversity [@], and zoonotic pandemics [@covid].
But will ML approaches really help us tackle a changing planet?
Applications of ML in ecology have begun to illustrate the promise of two methods of ML: *supervised learning* (e.g. image recognition) and *unsupervised learning* (including clustering and prediction), but have so far largely overlooked the third and possibly most promising approach in the ML triad: Reinforcement Learning (RL).
Three features distinguish RL from other ML methods in ways that are particularly well suited to addressing global ecological change issues:

@@

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the uncertain and changing environments considered by RL research have largely been drawn from examples in robotic movement and video games.
RL agents have recently achieved superhuman performance on classic Atari games [@DQN], while Deepmind’s RL-trained AlphaZero has defeated the world's grandmasters in Go, Chess, and Shogi [@alphazero].
Complex environmental problems share many similarities to these games: the need to plan many moves ahead, accept short-term costs to achieve long-term benefit, to account for uncertainty and respond with contingency to the unexpected.
Of course real ecosystems are also far more complex than a game of chess or MoonLander -- but can the same be said of our efforts at finding the best strategy?
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or global circulation models (GCMs) in climate [@gcm] are already used to explore scenarios and inform ecosystem management. 
Even simpler models are frequently employed throughout ecology to guide understanding and response.
<!-- Kari: ick, jargon heavy.  Need to say: some existing methods, but limited to simple problems or simple solutions -->
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of models, the so called "curse of dimensionality" makes problems with too many possible states or actions intractable by conventional methods such as stochastic dynamic programming (SDP) and approximate dynamic programming (ADP) [@Polasky2011; @Marescot2013].
Neural-network-based ML techniques have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.
We believe that the combination of such approaches from RL trained using sufficient simulations of suitable ecological processes will open the door to a new paradigm for tackling tough environmental change problems.
We also believe that such a transformation could bring risks and pitfalls comparable and varied as its solutions.

@@

@@

@@

Meanwhile, the state-action-value or quality function, finds the expected return given a state and action, $$
  Q^\pi(s_t, a_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t, a_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

@@

Should we mention SARSA, Double DQN?  
Can we add HER (since delayed reward case may be particularly interesting)
Should we do anything to better reflect the 'family tree' here? (e.g. DQN -> DDPG -> TD3)
